---
title: "Reading data from website"
output: github_document
---

## Two major paths
1. there is data included as content on a webpage, and you want to 'scrape' those data
- table from Wikipedia
- reviews from Amazon
- cast and characters on IMBD

2. there is a dedicated server holding data in a relatively usable form, and you want yo ask for those data
- open NYC data
- Data.gov
- Star Wars API (Application Programming Interfaces)


```{r setup, include = FALSE}
#always to include, default set

library(tidyverse)

library(rvest)
library(httr)


theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "virids",
  ggplot2.continuous.fill = "virids"
)

scale_color_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```

# scape table
workfolw for scraping data

- download html using `read_html()`
- extract nodes using `html_notes()` and your CSS selector
- extract content from nodes using `html_text()`,`html_table()`, etc


I want first table from [this page] (https://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm)

read in the html
```{r}
url = "https://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

drug_use_html = read_html (url)
```

extract the table(s); focus on the first one
```{r}

table_marj <- html_nodes(drug_use_html, css = "table")[[1]] |> #use [[1]] extract the first item
  html_table() |> # to gengerate a table to read
  slice(-1) |> #elimate first row "notes" 
  as_tibble() #convert to table
```


## star war movies info

I want the data from [here] (https://www.imdb.com/list/ls070150896/).

```{r}
url = "https://www.imdb.com/list/ls070150896/"

swm_html = read_html(url)
```

grab elements that I want.
```{r}
title_vec =
  swm_html |>
  html_nodes(css = ".lister-item-header a")|>
  html_text()


gross_rev_vec = 
  swm_html |>
  html_nodes(css = ".text-small:nth-child(7) span:nth-child")|>
  html_text()

runtime_vec =
  swm_html |>
  html_nodes(css = ".runtime")|>
  html_text()
  
swm_df = 
  tibble(
    title = title_vec,
    gross_rec = gross_rev_vec,
    runtime = runtime_vec)
```

# get some water data 
this is coming from an API

```{r}
nyc_water =
  GET("https://data.cityofnewyork.us/resource/ia2d-e54m.csv") |>
  content("porsed")


nyc_water_json =
  GET("https://data.cityofnewyork.us/resource/ia2d-e54m.json") |>
  content("text")|>
  jsonlite::fromJSON()|>
  as_tibble()
```


# BRFSS

same process, different data
```{r}
brfss_2010 = 
  GET("https://chronicdata.cdc.gov/resource/acme-vg9e.csv",
      query = list("$limit" = 5000))|>
  content("parsed")

```






