---
title: "Strongs and factors"
output: github_document
---

## Two major paths
1. there is data included as content on a webpage, and you want to 'scrape' those data
- table from Wikipedia
- reviews from Amazon
- cast and characters on IMBD

2. there is a dedicated server holding data in a relatively usable form, and you want yo ask for those data
- open NYC data
- Data.gov
- Star Wars API (Application Programming Interfaces)


```{r setup, include = FALSE}
#always to include, default set

library(tidyverse)
library(p8105.datasets)
library(rvest)
library(httr)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "virids",
  ggplot2.continuous.fill = "virids"
)

scale_color_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Strings and regex

```{r}
string_vec = c("my", "name", "is", "jeff")

#`str_etect` the presence or absence of a pattern in a string
str_detect(string_vec, "jeff") #does "jeff" exist in the string
#[1] FALSE FALSE FALSE  TRUE


#replace pattern
str_replace(string_vec, "jeff", "Jeff")
```


```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )

#start with "i think" by using ^ at the beginning
str_detect(string_vec, "^i think")

#end with "i think" by using $ at the end
str_detect(string_vec, "i think$")
```

```{r}
string_vec = c(
  "Y'all remember Pres. HW Bush?",
  "I saw a green bush",
  "BBQ and Bushwalking at Molonglo Gorge",
  "BUSH -- LIVE IN CONCERT!!"
)

#to detect both "Bush" and "bush"
str_detect(string_vec, "[Bb]ush")
#[1]  TRUE  TRUE  TRUE FALSE
```


```{r}
string_vec = c(
  "Time for a Pumpkin Spice Latte!",
  "went to the #pumpkinpatch last weekend",
  "Pumpkin Pie is obviously the best pie",
  "SMASHING PUMPKINS -- LIVE IN CONCERT!!"
  )

str_detect(string_vec,"[Pp]umpkin")

```

```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

#want to find number first, immediately follow by characters
str_detect(string_vec, "^[0-9][a-zA-Z]")
#[1]  TRUE  TRUE FALSE  TRUE
```

```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

#`^` target any special character at the beginning;
#`.` target any special character in the middle;
#`$` target ant special character at the end;
str_detect(string_vec, "7.11")
#[1]  TRUE  TRUE FALSE  TRUE

#if we want to target only `7.11`, dectect actual dot `.`
str_detect(string_vec, "7\\.11")

```

```{r}
string_vec = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]'
  )

# to detect specific special character, using "\\"
str_detect(string_vec, "\\[")
#[1]  TRUE FALSE  TRUE  TRUE

```

## Factors

```{r}
factor_vec = factor(c("male","male","female","female"))
factor_vec
#[1] male   male   female female
#Levels: female male (based on alphabet)

as.numeric(factor_vec)
#[1] 2 2 1 1; female = 2, male = 1 

```

what happen if i relevel ..

```{r}
factor_vec = fct_relevel(factor_vec, "male")
factor_vec
#[1] male   male   female female
#Levels: male female

as.numeric(factor_vec)
#[1] 1 1 2 2; male = 1, female = 2

```

## NSDUH -- Strings
read in the html
```{r}
url = "https://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

drug_use_html = read_html (url)
```

extract the table(s); focus on the first one
```{r}
table_marj <- html_nodes(drug_use_html, css = "table")[[1]] |> #use [[1]] extract the first item
  html_table() |> # to gengerate a table to read
  slice(-1) |> #elimate first row "notes" 
  as_tibble() #convert to table
```


```{r}
data_marj <- 
  table_marj |>
  select(-contains("P Value")) |>   # Select all columns except those with "P Value"
  pivot_longer(
    cols = -State,                  # Specify columns to keep in long format
    names_to = "age_year",
    values_to = "percent"           # Specify the output column name for values
  ) |> 
  separate(age_year, into = c("age", "year"), sep = "\\(") |> 
  mutate(
    year = str_replace(year, "\\)", ""),      # Remove the closing parenthesis from `year`
    percent = str_replace(percent, "[a-c]$", ""),  # Remove specific trailing characters from `percent`
    percent = as.numeric(percent)             # Convert `percent` to numeric
  ) |> 
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West"))) # Filter out specific regions

```


## NSDUH -- Factors

```{r}
data_marj |>
  filter(age == "12-17") |> 
  mutate(State = fct_reorder(State, percent))|> #put state in order with percentage
  ggplot(aes(x = State, y = percent, color = year)) +  
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```


















